Network based multifactorial modelling microRNA-target interaction.

### 2019-05-28 article.tex
todo-tofix

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Yildiz Technical University Department of
  Bioengineering} \\ Alper Yilmaz\footnote{Yildiz Technical University Department of
  Bioengineering}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{17 May 2019}

@@

Competing endogeneous RNA (ceRNA) regulations and crosstalk in human are
remarkable issue in mean of regulations via miRNAs. Many studies have
pointed out that an alteration in miRNA:target interaction can result in
unexpected regulations because of indirect and complex interaction. In
this paper, we defined a new network-based model that handles
miRNA:ceRNA interactions with expression values and interaction factors
optionally. We applied approach in sample dataset and defined dynamics
of it. We have observed that a perturbation in an element of network
affects whole competing elements differently and cooperative
efficiencies of miRNAs on common targets could be shown. Importance of
miRNA:target ratios was also observed consistently with previous
reports. We have obtained that the competing elements which have the
same or close expression values may not affected equal from the
perturbation because of repression functionality of interactions. To
apply model to the real sample that obtained combining breeast cancer
gene and miRNA expression dataset via high throughput experimental
miRNA:target dataset, we selected a gene, \emph{ABCC1}, which over
expressed in tumor tissue and utilized it as perturbating element. We
have observed that one change can be sufficient to perturb regulations,
and may not predicted all of regulations correctly, with this change.
Therefore, this model which helps to understand crosstalking between
elements could be tested and improved via results of well-constructed
experimental dataset and takes to hope for future research.

@@

MicroRNAs (miRNAs) that are key regulator of gene expression through
various post-transcriptional mechanisms are a family of short non-coding
RNAs. Although repression mechanisms of miRNA are not fully understood,
exact effect of miRNAs is regulation of direct target/targets
negatively. Repressive activities of miRNAs vary depending on many
factors that are significant to microRNA:target interactions. These
factors include microRNA:target binding energy, binding location in
target sequence, base pairing types between miRNA and target, abundance
of miRNAs and targets (Grimson et al. 2007). Binding energies of
miRNA:target complexes vary based on nucleotide context and determine
folding stability of complex (Cao and Chen 2012). It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex (Helwak et al. 2013) and does not
directly determine repressive activity of miRNA (Cao and Chen 2012).
Early studies have argued that 2-8 nt sequence, seed, located in miRNA
5'UTR bind to specific sequence located in 3'UTR of its target (Bartel
2004; Lewis, Burge, and Bartel 2005). In recent studies, it has been
shown miRNAs can interact with targets via sequence that located in
different region such as 5'UTR or CDS (Hausser et al. 2013; Helwak et
al. 2013) \textbf{moore\_mirnatarget\_2015} (Moore et al. 2015). These
studies also shown that binding location could indicate functionality of
miRNA:target interaction or affect profile of target abundance. It has
been shown that miRNAs exhibit repressive activity via, 6-8 nt long
sequence that is prefectly complementary with targets, seed region at
the 5'UTR of miRNAs (Bartel 2009; Grimson et al. 2007). On the other
hand, some researchers have reported that seed sequence of miRNA can
have mismatches or bulged/wobble nucleotides and locates in different
region of miRNAs (Hafner et al. 2010; Helwak et al. 2013). In addition
to all of these, abundance of miRNAs and targets and miRNA:target ratio
in cells are determinant for efficiency of miRNA:target interaction
(Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al.
2014).

As it is possible for miRNAs to suppress multiple targets, an individual
RNA molecule can also be targeted by various miRNAs. In that case, the
targeted RNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
2013), against their miRNAs. Briefly, Ala et al.~have explained the
ceRNA hypothesis as disturbtion of the other target when a steady-state
system that included one miRNA and two target, was perturbed with
expression changes at one of the targets (Ala et al. 2013). When
interaction between miRNAs and their target were dealed in the cell,
explanation and prediction of results of perturbation is difficult due
to varieties of interactions. Various computational and experimental
studies have applied to be understood for ceRNA:miRNA interactions. It
has been observed that when abundance of one of the targets of miR-122
was increased, the other target expression also slightly increased as a
result of decreasing repressive activity of miR-122 on them (Denzler et
al. 2014). Bosson et al.~have developed a mathematical model for changes
on total target pool concentration after grouping to affinity of
targets. They have demonstrated that miRNA activity correlated with
affinity between miRNA and target (Bosson, Zamudio, and Sharp 2014).
Cooperative efficiency of miRNAs as well as competitor behaviors of
targets also were studied and it has been demonstrated crucial for
regulations at expression of targets (Denzler et al. 2016). MiRNA:target
interactions have modelled as stoichiometric and catalytic mechanisms
and Figliuzzi et al.~have recommended to handling models in network
structures (Figliuzzi, Marinari, and De Martino 2013). The model that
can explain miRNA target interaction through topological features has
been applied at bipartite network by Nitzan et al.~(Nitzan et al. 2014).
Robinson and Henderson applied the model that handles miRNA:target
direct and indirect interactions via common miRNA of gene and target of
miRNA, at bipartite network. It has been demonstrated that whole miRNAs
and targets in network can interacted through common miRNAs and genes
(Robinson and Henderson 2018). Associated genes that are targets of the
same miRNAs have been found with help of correlation of gene expression
changes in recent study (Markus List 2017). List et al.~have specified
that their approach can be useful for ceRNA's studies and published
these approach as a package that can use with R programming.

@@

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Shematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: triangle grey) repress targets (G: circle
coconut) according to proportion of their targets' expression. b-c-d)
MiRNA expression distributions to targets vary when expression of one of
the targets perturbed.}\label{fig1}
}
\end{figure}

The model was developed by making some assumptions. Transcription rates
and degradation rates of miRNAs were assumed as stationary and equal.
Targets also have stabil transcription and degradation rates, at the
same time they are equal. The factors that could be affect the
miRNA:target interaction are disregarded in this approach. It is
considered that miRNAs degradates or inhibits them via any mechanisms
after binding to target while the miRNAs were used reversibly. In
initial conditions, gene expression amounts in system have been
considered after repression activity of miRNAs. If a change occurs in
the system, variation on miRNA proportional distribution are reflected
as change of repression activity. \begin{equation} 
    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
\end{equation} \begin{equation}
   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

@@

To determine suitability of method, we have tried to apply model in a
real dataset. Firstly, we obtained the miRNA and gene expression
datasets of the patient's tumor and normal tissue. We downloaded the
high-throughtput experimental datasets which are provided miRNA:gene
target pairs with interaction factors by
\textbf{moore\_mirnatarget\_2015} and (Helwak et al. 2013). We have
combined miRNA and gene expression datasets via miRNA:target gene
dataset \textbf{(Whole steps can be found in Supplementary
data(TCGA\_E9-A1N5\_article.Rmd)}). We selected a gene which is over
expressed in tumor tissue with compared to normal tissue. It is pointed
out that ABCC1 gene, \emph{Multidrug resistance-associated protein 1},
is one of the most significant factor to develop resistance against
chemotherapotic agents(Atalay, Demirkazik, and Gunduz 2008; Lu et al.
2015; Atalay et al. 2006). So, we selected ABCC1 gene for simulation of
integrated dataset. We have detected the iteration that is suitable for
ABCC1 gene simulation. After simulation of system via ABCC1 gene with
over expression, we have compared simulation results and tumor tissue
expression levels.

@@

We have developed a model approach and workflow for competitive ceRNA
regulation in this study. The basic mode of miRNA repression activity
has been based on miRNA and target abundance in various researches(Arvey
et al. 2010; Denzler et al. 2014). For this reason, we performed first
application as considering only abundance of miRNAs and targets
(\autoref{fig1}). After the increasing in a gene expression value,
expression values of other genes also changed slightly in comparison
with alteration in perturbation started gene node. Different studies
have shown that the other gene expression values also rise differently,
if a gene abundance increases in ceRNA system (Lai, Wolkenhauer, and
Vera 2016; Salmena et al. 2011; Tay, Rinn, and Pandolfi 2014). It was
observed that primary neighborhoods of the trigger gene change with the
same ratio, but cooperative efficiency of miRNAs on common target causes
differently change in common target (\autoref{fig1}d). The common target
in system has displayed like a trigger for the other group and induced
changes of expression values of genes on the other group. Therefore,
effect of primary change of the gene (G2) on genes of the other group
was observed slightly (\autoref{fig1}d). In addition, like shown by
ceRNA hypothesis model of Ala et al., after the increasing of a gene
expression (G2), the miRNA that is found in the same group (M1) tended
to be more repressive on this target. This has caused to decrease in
increased expression value of primer triggering gene. When the
expression of a gene (G2) was increased to two fold, we observed that
the expression changes were more evident. So, we have considered that
high miRNA:target ratio results in stronger miRNA repression activity.
This results also are coherent with previous reports that were offered
by Arvey et al.~and Denzler et al.(Arvey et al. 2010; Bosson, Zamudio,
and Sharp 2014; Denzler et al. 2014). When the simulation was run in
sample dataset that includes lower target abundance, the first response
of primary neihgborhood of the same group was determined as similar with
in system that have high target abudance. But regulations of ceRNAs were
observed differences in the folowing steps of simulation. In overall
state in the end of simulation, we have established that all regulations
more prominent when compared with low miRNA:target ratio dataset. The
model has appeared the importance of mRNA:target ratio in regulation of
competing RNAs.

@@

It has been reported in several studies that miRNA regulatory
interactions also is affected different parameters. For example, Xu et.
al have investigated the importance of seed pairing type between miRNAs
and their targets and target site location by using proteomic dataset
(Xu, Wang, and Liu 2014). They have proposed that the features of
binding between miRNA and target can be critical efficiency of miRNA and
can be used to find effective miRNA molecules. In addition, binding
energy between miRNA and targets also could be determinant for miRNA
efficiency. It has been reported by Breda et. al.~that strength of
miRNA:target interactions is depended on binding energy of complexes
(Breda et al. 2015). As similar to these, Bosson et al.~have associated
affinity and seed pairing of miRNA target pairs and suggested that
number of canonical base pairing is correlated with affinity. For these
reasons, we have improved the model with optional interaction parameters
that could be useful for understanding of miRNA repression activity. We
have explained the model in the same sample dataset with using different
factors and simulated via the same change (\autoref{fig3}). When the
factors were taken into account in the system, miRNA efficiencies varied
as shown in \autoref{fig3}a. Although the miRNA target ratios in initial
states were same in comparison with the sample dataset without factors,
efficiency of binding and repression have changed. When expression of a
gene (G2) increased, expression values of all genes also change
differentially because of contribution of efficiency factors
(\autoref{fig3} b,c,d). We have considered that in our approach energy
and seed type of pairs is significant for binding and targeted region is
important for repression. But these factors can be handled optionally
and differently.

@@

The defined miRNA' number is very low in comparison with defined protein
coding genes' number. So, this results in that the miRNA amounts in the
system must be high or targeted RNAs' amount must be low for effective
miRNA activity. We have tried to simulate a real dataset which contains
thousands of genes and hundreds of miRNAs. Firstly, we have observed the
activities in the iterations and determined behaviours of the network
after simulation. As expected, we have observed that the big dataset,
compared to those containing a few genes and miRNAs, exhibited
cumbersome behaviours and complexity to gain steady-state \textbf{(see
details in Supplementary File(TCGA\_E9-A1N5\_article.Rmd)}. We realized
that a small change in a system with a large change actually brought
many arrangements. This can be explained with the competing groups.
Because the genes are targeted with many common miRNAs, and this results
in that the common targets exhibit competing behaviours in different
groups. On the other hand, the changes of gene expressions were very
different. While the change in some of genes were highly evident, somes
have changed slightly. So, we have compared two datasets; the tumor gene
expression and simulation results. When we analysed comparison of two
datasets, we have observed that the results of simulation is not
convenient with tumor tissue expression values. The simulation results
were not expected to be concordant with tumor tissue expression values
because of a single gene could not be responsible for gene regulation in
all tissue. Besides, the all other factors such as up/down regulation of
miRNAs or other genes were ignored. Of course, it would be more useful
to test a dataset that includes gene expression values before and after
the regulation of a gene and miRNA expression values at initial
conditions. When we evaluated the suitability of approach, the approach
allows different miRNA targets to be considered together and can work
with different regulatory elements due to the network structure.

The network based approaches have been developed in previous studies.
Figliuzzi et al.~have tried to explain the ceRNA crosstalk in a
network-like minimal interaction structure with concentrations of ceRNA
and miRNAs. They have pointed out that the larger miRNA number can cause
to evident crosstalk between ceRNAs. It has been thought that miRNAs and
their targets interactions depend on rates of transcription,
degradation, binding and unbinding in the network based kinetic model
which has been developed by Nitzan et al.~Nitzan and collaborators have
used the high throughput experimental dataset (Helwak et al. 2013) for
miRNA:target pair that occur high free energy and microarray dataset of
a miR-92a depletion for expression values. They have been demonstrated
that distant ceRNAs can interact with each other via indirect links, and
the interactions change to depend on distance between ceRNAs, and
topological features of network (Nitzan et al. 2014). List et al.~have
developed an approach to detect ceRNA interaction by using the miRNA
expression, gene expression and common miRNAs between gene targets
(Markus List 2017). Their approach could be useful to analysis
interacted genes through miRNAs. When it has been throught that a miRNA
can exhibit strong functionality to a target but may not against an
other, common miRNA based approach can be inconvenient to understand
regulations of ceRNA interactions.

@@

\leavevmode\hypertarget{ref-figliuzzi_micrornas_2013}{}%
Figliuzzi, Matteo, Enzo Marinari, and De MartinoAndrea. 2013.
``MicroRNAs as a Selective Channel of Communication Between Competing
RNAs: A Steady-State Theory.'' \emph{Biophysical Journal} 104 (5):
1203--13. \url{https://doi.org/10.1016/j.bpj.2013.01.012}.

>>>>

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Yildiz Technical University Department of
  Bioengineering} \\ Alper Yilmaz\footnote{Yildiz Technical University Department of
  Bioengineering}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{28 Mayıs 2019}

@@

Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı
ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
We applied the model to real sample consisting of breast cancer gene and
miRNA expression dataset and experimental miRNA:target interaction
dataset all generated via high throughput sequencing methods. A gene
over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
perturbing element. We have observed that change in expression level of
single gene in miRNA:target network is sufficient to perturb regulations
in whole network, due to unforeseen and unpredicted regulation which are
only visible when considered in network context. Therefore, this model
helps unveiling the crosstalk between elements in miRNA:target network
where abundance of target and sponge effect are taken into account. The
model is scalable and can be plugged in with emerging miRNA effectors
such as circRNAs. The model is available as R package at
\url{https://github.com/selcenari/ceRNAnetsim}. (TODO the link will
change)

@@

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to microRNA:target interactions. These factors include microRNA:target
binding energy, binding location in target sequence, base pairing types
between miRNA and target, abundance of miRNAs and targets (Grimson et
al. 2007). Binding energies of miRNA:target complexes vary based on
nucleotide context and determine folding stability of complex (Cao and
Chen 2012). It has been demonstrated that the binding energy between
miRNA and target indicates stability or affinity of complex (Helwak et
al. 2013) and does not directly determine repressive activity of miRNA
(Cao and Chen 2012). Early studies have argued that 2-8 nt sequence,
seed, located in miRNA 5'end bind to specific sequence located in 3'UTR
of its target (Bartel 2004; Lewis, Burge, and Bartel 2005). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al. 2013;
Helwak et al. 2013; Moore et al. 2015). These studies also showed that
binding location could indicate functionality of miRNA:target
interaction or be effective on abundance of targets(TOFIX abundance
etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?,
evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that
miRNAs exhibit repressive activity via, 6-8 nt long sequence that is
perfectly complementary with targets, seed region at the 5' end of
miRNAs (Bartel 2009; Grimson et al. 2007). On the other hand, some
researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides and may locate in region other
than 5'end of miRNAs (Hafner et al. 2010; Helwak et al. 2013). On top of
all these factors, abundance of miRNAs and targets and miRNA:target
ratio in cells predominantly affect efficiency of miRNA:target
interaction (Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler
et al. 2014).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
2013), against their miRNAs. Briefly, Ala et al.~have explained the
ceRNA hypothesis as disturbance of the other target when one of the
targets on a steady-state system that included one miRNA and two target
was perturbed with expression change (Ala et al. 2013). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other target expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al. 2014). Bosson et al.~have developed a mathematical
model for changes on total target pool concentration after grouping
targets according to affinity and demonstrated that miRNA activity
correlated with affinity between miRNA and target (Bosson, Zamudio, and
Sharp 2014). Cooperative efficiency of miRNAs as well as competitor
behaviors of targets were also studied and it has been demonstrated to
be crucial for regulating available mRNA levels of targets (Denzler et
al. 2016). MiRNA:target interactions have been modeled as stoichiometric
and catalytic mechanisms and Figliuzzi et al.~have recommended handling
models in network context (Figliuzzi, Marinari, and De Martino 2013).
The model that can explain miRNA target interaction through topological
features has been applied at bipartite network by Nitzan et al.~(Nitzan
et al. 2014). Robinson and Henderson applied the model that handles
miRNA:target direct and indirect interactions via common miRNA of gene
and target of miRNA, at bipartite network. It has been demonstrated that
all miRNAs and targets in the network can interact with each other
through common miRNAs and genes, without interaction between the same
type of nodes (Robinson and Henderson 2018).Associated genes that are
targets of the same miRNAs have been found with help of correlation of
gene expression changes in recent algoritm (Markus List 2017). List et
al.~have specified that their approach can be useful for ceRNA studies
and published their approach as an R package.

@@

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Schematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: gray triangle) repress targets (G: circle
coconut) according to proportion of their targets' expression. b-c-d)
MiRNA expression distributions to targets vary when expression of one of
the targets perturbed.}\label{fig1}
}
\end{figure}

The model was developed by making some assumptions. Transcription rates
and degradation rates of miRNAs were assumed as stationary and equal.
Targets also have stable transcription and degradation rates, at the
same time they are equal. The factors that could be affect the
miRNA:target interaction are disregarded in this approach. It is
considered that miRNAs degrades or inhibits them via any mechanisms
after binding to target while the miRNAs were used reversibly. In
initial conditions, gene expression amounts in system have been
considered after repression activity of miRNAs. If a change occurs in
the system, variation on miRNA proportional distribution are reflected
as change of repression activity. \begin{equation} 
    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
\end{equation} \begin{equation}
   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

@@

To determine suitability of method, we have tried to apply model in a
real dataset. Firstly, we obtained the miRNA and gene expression
datasets of the patient's tumor and normal tissue. We downloaded the
high-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors by
\textbf{moore\_mirnatarget\_2015} and (Helwak et al. 2013). We have
combined miRNA and gene expression datasets via miRNA:target gene
dataset \textbf{(Whole steps can be found in Supplementary
data(TCGA\_E9-A1N5\_article.Rmd)}). We selected a gene which is over
expressed in tumor tissue with compared to normal tissue. It is pointed
out that ABCC1 gene, \emph{Multidrug resistance-associated protein 1},
is one of the most significant factor to develop resistance against
chemotherapotic agents(Atalay, Demirkazik, and Gunduz 2008; Lu et al.
2015; Atalay et al. 2006). So, we selected ABCC1 gene for simulation of
integrated dataset. We have detected the iteration that is suitable for
ABCC1 gene simulation. After simulation of system via ABCC1 gene with
over expression, we have compared simulation results and tumor tissue
expression levels.

@@

We have developed a model approach and workflow for competitive ceRNA
regulation in this study. The basic mode of miRNA repression activity
has been based on miRNA and target abundance in various researches(Arvey
et al. 2010; Denzler et al. 2014). For this reason, we performed first
application as considering only abundance of miRNAs and targets
(\autoref{fig1}). After the increasing in a gene expression value,
expression values of other genes also changed slightly in comparison
with alteration in perturbation started gene node. Different studies
have shown that the other gene expression values also rise differently,
if a gene abundance increases in ceRNA system (Lai, Wolkenhauer, and
Vera 2016; Salmena et al. 2011; Tay, Rinn, and Pandolfi 2014). It was
observed that primary neighborhoods of the trigger gene change with the
same ratio, but cooperative efficiency of miRNAs on common target causes
differently change in common target (\autoref{fig1}d). The common target
in system has displayed like a trigger for the other group and induced
changes of expression values of genes on the other group. Therefore,
effect of primary change of the gene (G2) on genes of the other group
was observed slightly (\autoref{fig1}d). In addition, like shown by
ceRNA hypothesis model of Ala et al., after the increasing of a gene
expression (G2), the miRNA that is found in the same group (M1) tended
to be more repressive on this target. This has caused to decrease in
increased expression value of primer triggering gene. When the
expression of a gene (G2) was increased to two fold, we observed that
the expression changes were more evident. So, we have considered that
high miRNA:target ratio results in stronger miRNA repression activity.
This results also are coherent with previous reports that were offered
by Arvey et al.~and Denzler et al.(Arvey et al. 2010; Bosson, Zamudio,
and Sharp 2014; Denzler et al. 2014). When the simulation was run in
sample dataset that includes lower target abundance, the first response
of primary neighborhood of the same group was determined as similar with
in system that have high target abundance. But regulations of ceRNAs
were observed differences in the following steps of simulation. In
overall state in the end of simulation, we have established that all
regulations more prominent when compared with low miRNA:target ratio
dataset. The model has appeared the importance of mRNA:target ratio in
regulation of competing RNAs.

@@

It has been reported in several studies that miRNA regulatory
interactions also is affected different parameters. For example, Xu et.
al have investigated the importance of seed pairing type between miRNAs
and their targets and target site location by using proteomics dataset
(Xu, Wang, and Liu 2014). They have proposed that the features of
binding between miRNA and target can be critical efficiency of miRNA and
can be used to find effective miRNA molecules. In addition, binding
energy between miRNA and targets also could be determinant for miRNA
efficiency. It has been reported by Breda et. al.~that strength of
miRNA:target interactions is depended on binding energy of complexes
(Breda et al. 2015). As similar to these, Bosson et al.~have associated
affinity and seed pairing of miRNA target pairs and suggested that
number of canonical base pairing is correlated with affinity. For these
reasons, we have improved the model with optional interaction parameters
that could be useful for understanding of miRNA repression activity. We
have explained the model in the same sample dataset with using different
factors and simulated via the same change (\autoref{fig3}). When the
factors were taken into account in the system, miRNA efficiencies varied
as shown in \autoref{fig3}a. Although the miRNA target ratios in initial
states were same in comparison with the sample dataset without factors,
efficiency of binding and repression have changed. When expression of a
gene (G2) increased, expression values of all genes also change
differentially because of contribution of efficiency factors
(\autoref{fig3} b,c,d). We have considered that in our approach energy
and seed type of pairs is significant for binding and targeted region is
important for repression. But these factors can be handled optionally
and differently.

@@

The defined miRNA' number is very low in comparison with defined protein
coding genes' number. So, this results in that the miRNA amounts in the
system must be high or targeted RNAs' amount must be low for effective
miRNA activity. We have tried to simulate a real dataset which contains
thousands of genes and hundreds of miRNAs. Firstly, we have observed the
activities in the iterations and determined behaviors of the network
after simulation. As expected, we have observed that the big dataset,
compared to those containing a few genes and miRNAs, exhibited
cumbersome behaviors and complexity to gain steady-state \textbf{(see
details in Supplementary File(TCGA\_E9-A1N5\_article.Rmd)}. We realized
that a small change in a system with a large change actually brought
many arrangements. This can be explained with the competing groups.
Because the genes are targeted with many common miRNAs, and this results
in that the common targets exhibit competing behaviors in different
groups. On the other hand, the changes of gene expressions were very
different. While the change in some of genes were highly evident, some
have changed slightly. So, we have compared two datasets; the tumor gene
expression and simulation results. When we analyzed comparison of two
datasets, we have observed that the results of simulation is not
convenient with tumor tissue expression values. The simulation results
were not expected to be concordant with tumor tissue expression values
because of a single gene could not be responsible for gene regulation in
all tissue. Besides, the all other factors such as up/down regulation of
miRNAs or other genes were ignored. Of course, it would be more useful
to test a dataset that includes gene expression values before and after
the regulation of a gene and miRNA expression values at initial
conditions. When we evaluated the suitability of approach, the approach
allows different miRNA targets to be considered together and can work
with different regulatory elements due to the network structure.

The network based approaches have been developed in previous studies.
Figliuzzi et al.~have tried to explain the ceRNA crosstalk in a
network-like minimal interaction structure with concentrations of ceRNA
and miRNAs. They have pointed out that the larger miRNA number can cause
to evident crosstalk between ceRNAs. It has been thought that miRNAs and
their targets interactions depend on rates of transcription,
degradation, binding and unbinding in the network based kinetic model
which has been developed by Nitzan et al.~Nitzan and collaborators have
used the high throughput experimental dataset (Helwak et al. 2013) for
miRNA:target pair that occur high free energy and microarray dataset of
a miR-92a depletion for expression values. They have been demonstrated
that distant ceRNAs can interact with each other via indirect links, and
the interactions change to depend on distance between ceRNAs, and
topological features of network (Nitzan et al. 2014). List et al.~have
developed an approach to detect ceRNA interaction by using the miRNA
expression, gene expression and common miRNAs between gene targets
(Markus List 2017). Their approach could be useful to analysis
interacted genes through miRNAs. When it has been thought that a miRNA
can exhibit strong functionality to a target but may not against an
other, common miRNA based approach can be inconvenient to understand
regulations of ceRNA interactions.

@@

\leavevmode\hypertarget{ref-figliuzzi_micrornas_2013}{}%
Figliuzzi, Matteo, Enzo Marinari, and Andrea De Martino. 2013.
``MicroRNAs as a Selective Channel of Communication Between Competing
RNAs: A Steady-State Theory.'' \emph{Biophysical Journal} 104 (5):
1203--13. \url{https://doi.org/10.1016/j.bpj.2013.01.012}.
